---
title: "Statistical inference course project"
author: "Mahesh"
date: "10 September 2016"
output: html_document
---
## Overview

Analyze the ToothGrowth data in the R datasets package.  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
data(ToothGrowth)
str(ToothGrowth)
```

### inspect the data

```{r}
head(ToothGrowth)
```

### Basic Summary

```{r}
summary(ToothGrowth)

#Unique Values
unique(ToothGrowth$len)

unique(ToothGrowth$supp)

unique(ToothGrowth$dose)
```


### Analysis
Let us plot some graphs to start exploratory analysis  

```{r}
# Convert dose to a factor
ToothGrowth$dose<-as.factor(ToothGrowth$dose)

# Plot tooth length ('len') vs. the dose amount ('dose'), broken out by supplement delivery method ('supp')
ggplot(aes(x=dose, y=len), data=ToothGrowth) + geom_boxplot(aes(fill=dose)) + xlab("Dose Amount") + ylab("Tooth Length") + facet_grid(~ supp) + ggtitle("Tooth Length vs. Dose Amount \nby Delivery Method") + 
     theme(plot.title = element_text(lineheight=.8, face="bold"))


# Plot tooth length ('len') vs. supplement delivery method ('supp') broken out by the dose amount ('dose')
ggplot(aes(x=supp, y=len), data=ToothGrowth) + geom_boxplot(aes(fill=supp)) + xlab("Supplement Delivery") + ylab("Tooth Length") + facet_grid(~ dose) + ggtitle("Tooth Length vs. Delivery Method \nby Dose Amount") + 
     theme(plot.title = element_text(lineheight=.8, face="bold"))
```

Use t test to compare growth by supplement  

```{r}
# run t-test
t.test(len~supp,data=ToothGrowth)

```

The p-value of this test was 0.06.  
Since the p-value is greater than 0.05 and the confidence interval of the test 
contains zero we can say that supplement types seems to have no impact on Tooth 
growth based on this test.  

Now we’ll compare tooth growth by dose, looking at the different pairs of 
dose values  

```{r}
# run t-test using dose amounts 0.5 and 1.0
ToothGrowth_sub <- subset(ToothGrowth, ToothGrowth$dose %in% c(1.0,0.5))
t.test(len~dose,data=ToothGrowth_sub)

# run t-test using dose amounts 0.5 and 2.0
ToothGrowth_sub <- subset(ToothGrowth, ToothGrowth$dose %in% c(0.5,2.0))
t.test(len~dose,data=ToothGrowth_sub)

# run t-test using dose amounts 1.0 and 2.0
ToothGrowth_sub <- subset(ToothGrowth, ToothGrowth$dose %in% c(1.0,2.0))
t.test(len~dose,data=ToothGrowth_sub)
```

the p-value of each test was essentially zero and the confidence interval of each test does not cross over zero (0).  

Based on this result we can assume that the average tooth length increases with an inceasing dose, and therefore the null hypothesis can be rejected.

### Conclusions

Given the following assumptions:

- The sample is representative of the population
- The distribution of the sample means follows the Central Limit Theorem

In reviewing our t-test analysis from above, we can conclude that supplement delivery method has no effect on tooth growth/length, however increased dosages do result in increased tooth length.




