---
title: "Statistical inference course project"
author: "Mahesh"
date: "10 September 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations. 

## Simulation (Part 1)

Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials.

Simulating this with <code>rexp(n, lambda)</code>

```{r}
# set seed for reproducability
set.seed(50)


nosim <- 1000
n <- 40
lambda <- 0.2
# create a 1000-row matrix in which each row consists of
# 40 samples drawn from the exp distribution
simdata <- matrix(rexp(nosim * n, rate = lambda), nosim)
# take the mean for each row
row_means <- rowMeans(simdata)
```

Now compute the Mean and Variance
<code>th_mean</code> to represent Theoritical Mean
<code>sa_mean</code> to represent Sample Mean
<code>th_var</code> to represent Theoritical Variance
<code>sa_var</code> to represent Sample Variance

```{r}
th_mean <- 1/lambda
sa_mean <- mean(row_means)
th_var <- (1/lambda)^2/n
sa_var <- var(row_means)
```

Lets examine the theoritical and sample mean and variance

### Sample Mean versus Theoretical Mean

```{r}
print(paste("Theoretical mean: ", th_mean));
print(paste("Sample mean: ", round(sa_mean, 3)))
```

### Sample Variance versus Theoretical Variance

```{r}
print (paste("Theoretical variance: ", th_var));
print(paste("Sample Variance: ", round(sa_var, 3)))

```

### Distribution

```{r}
# plot the histogram of averages
hist(row_means, breaks=50, prob=TRUE,
     main="Distribution of averages of samples,
     drawn from exponential distribution with lambda=0.2",
     xlab="")
# density of the averages of samples
lines(density(row_means))
# theoretical center of distribution
abline(v=1/lambda, col="red")
# theoretical density of the averages of samples
xfit <- seq(min(row_means), max(row_means), length=100)
yfit <- dnorm(xfit, mean=1/lambda, sd=(1/lambda/sqrt(n)))
lines(xfit, yfit, pch=22, col="red", lty=2)
# add legend
legend('topright', c("simulation", "theoretical"), lty=c(1,2), col=c("black", "red"))
```